---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# knapsack

<!-- badges: start -->
<!-- badges: end -->

The knapsack problem is a discrete optimization problem where we have a knapsack that can take a limited weight W and we want to fill this knapsack with a number of items i = 1; :::; n, each with a weight wi and a value vi. The goal is to find the knapsack with the largest value of the elements added to the knapsack.  
This problem is NP-hard, meaning that it is ”at least as hard as the hardest problem in [NP](https://en.wikipedia.org/wiki/NP-hardness)”. NP is a (fundamental) class of problems for which there are (currently) no polynomial time algorithms to solve them. It is an open (Millennium Prize) problem, whether it is or is not possible to solve these problems in polynomial time. For a more detailed background of the knapsack problem see [this page](https://en.wikipedia.org/wiki/Knapsack_problem).

## Installation

You can install the development version of knapsack like so:

```r
install.packages("devtools")
library(devtools)
install_github("LindaRenJonsson/knapsack")
```

## Examples

### brute_force_knapsack

```{r brute_force_knapsack}
library(knapsack)

set.seed(42)
n <- 2000
knapsack_objects <- data.frame(w = sample(1:4000, size = n, replace = TRUE), v <- runif(n = n, 0, 10000))
brute_force_knapsack(x = knapsack_objects[1:8,], W = 3500)
```


### greedy_knapsack

```{r greedy_knapsack}
set.seed(42)
n <- 2000
knapsack_objects <- data.frame(w = sample(1:4000, size = n, replace = TRUE), v <- runif(n = n, 0, 10000))
greedy_knapsack(x = knapsack_objects[1:800,], W = 3500)
```

