---
title: "lab_report_knapsack"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{lab_report_knapsack}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(knapsack)
```

# knapsack

The knapsack problem is a discrete optimization problem where we have a knapsack that can take a limited weight W and we want to fill this knapsack with a number of items _i = 1,..., n_, each with a weight _w<sub>i</sub>_ and a value _v<sub>i</sub>_. The goal is to find the knapsack with the largest value of the elements added to the knapsack.  
This problem is NP-hard, meaning that it is ”at least as hard as the hardest problem in [NP](https://en.wikipedia.org/wiki/NP-hardness)”. NP is a (fundamental) class of problems for which there are (currently) no polynomial time algorithms to solve them. It is an open (Millennium Prize) problem, whether it is or is not possible to solve these problems in polynomial time. For a more detailed background of the knapsack problem see [this page](https://en.wikipedia.org/wiki/Knapsack_problem).

## Installation

You can install the development version of knapsack like so:

```r
install.packages("devtools")
library(devtools)
install_github("LindaRenJonsson/knapsack")
```

## Examples

### brute_force_knapsack

```{r brute_force_knapsack}
library(knapsack)

set.seed(42)
n <- 2000
knapsack_objects <- data.frame(w = sample(1:4000, size = n, replace = TRUE), v <- runif(n = n, 0, 10000))
brute_force_knapsack(x = knapsack_objects[1:8,], W = 3500)
```


### greedy_knapsack

```{r greedy_knapsack}
set.seed(42)
n <- 2000
knapsack_objects <- data.frame(w = sample(1:4000, size = n, replace = TRUE), v <- runif(n = n, 0, 10000))
greedy_knapsack(x = knapsack_objects[1:800,], W = 3500)
```


#### NOTE about assignment

Do ___note___ that the assignment gives answers for the `greedy_knapsack` that are not correct. This is easy to show using code:

```{r}
set.seed(42)
n <- 2000
knapsack_objects <- data.frame(w = sample(1:4000, size = n, replace = TRUE),
v <- runif(n = n, 0, 10000))

# greedy_knapsack(x = knapsack_objects[1:800,], W = 3500)

# Correct answers
# Value:
sum(knapsack_objects$v[c(256, 530, 701, 559,  89,  75, 759, 626, 219, 244,  63, 672, 455, 764, 329,  77, 705, 320, 110, 509, 762, 729, 691, 283, 553, 620, 341, 187,  83, 707, 511, 322)])

# Weight:
sum(knapsack_objects$w[c(256, 530, 701, 559,  89,  75, 759, 626, 219, 244,  63, 672, 455, 764, 329,  77, 705, 320, 110, 509, 762, 729, 691, 283, 553, 620, 341, 187,  83, 707, 511, 322)])


# Given answer is missing element on index 322
# Value:
sum(knapsack_objects$v[c(256, 530, 701, 559,  89,  75, 759, 626, 219, 244,  63, 672, 455, 764, 329,  77, 705, 320, 110, 509, 762, 729, 691, 283, 553, 620, 341, 187,  83, 707, 511)])

# Weight: 
sum(knapsack_objects$w[c(256, 530, 701, 559,  89,  75, 759, 626, 219, 244,  63, 672, 455, 764, 329,  77, 705, 320, 110, 509, 762, 729, 691, 283, 553, 620, 341, 187,  83, 707, 511)])


# greedy_knapsack(x = knapsack_objects[1:1200,], W = 2000)

# Correct answers

# Value:
sum(knapsack_objects$v[c(256,  530, 1186,  951,  701, 1157,  973,  559, 1019,   89,   75,  759,  845, 1153,  626,  897, 1106,  219, 1065, 1070,  244, 1036,  63,  672,  455,  826,  764,  981,  329,   77,  320,  110)])

# Weight:
sum(knapsack_objects$w[c(256,  530, 1186,  951,  701, 1157,  973,  559, 1019,   89,   75,  759,  845, 1153,  626,  897, 1106,  219, 1065, 1070,  244, 1036,  63,  672,  455,  826,  764,  981,  329,   77,  320,  110)])


# Given answer is missing elements on indices 320 and 110
# Value:
sum(knapsack_objects$v[c(256,  530, 1186,  951,  701, 1157,  973,  559, 1019,   89,   75,  759,  845, 1153,  626,  897, 1106,  219, 1065, 1070,  244, 1036,  63,  672,  455,  826,  764,  981,  329,   77)])

# Weight:
sum(knapsack_objects$w[c(256,  530, 1186,  951,  701, 1157,  973,  559, 1019,   89,   75,  759,  845, 1153,  626,  897, 1106,  219, 1065, 1070,  244, 1036,  63,  672,  455,  826,  764,  981,  329,   77)])
```
